$R^*$: A robust MCMC convergence diagnostic with uncertainty via gradient-boosted machines

### 2020-03-14 write_up/ms.tex
added more words to intro

Its importance in applied scientists' toolkits means it is essential the MCMC is used properly and with adequate care. A cost of automated inference software is that it is increasingly easy to regard MCMC as oracular: giving uncompromised depictions of the posterior. In spite of this ease-of-application, many statisticians who develop MCMC methods are less at ease with recommending their use. We repeat here a quote from Chris Holmes to this effect,

@@

Because of this, software packages (Stan in particular \cite{carpenter2017stan} is a great exemplar of this), go to great lengths to communicate to users any issues with sampling.

>>>>

Its importance in applied scientists' toolkits means it is essential that MCMC is used properly and with adequate care. A cost of automated inference software is that it is increasingly easy to regard MCMC as oracular: giving uncompromised depictions of the posterior. Because of this, software packages (Stan in particular \cite{carpenter2017stan} is a great exemplar of this), go to great lengths to communicate to users any issues with sampling. Indeed, in spite of this ease-of-application, many statisticians who develop MCMC methods are less at ease with recommending their use. We repeat here a quote from Chris Holmes to this effect \cite{holmes},

@@

The most important determination of whether MCMC has worked is whether the sampling distribution has converged to the posterior \cite{lambert2018Student}. MCMC methods are thus created because of an asymptotic property: that given an infinite number of samples, their sampling distribution approaches the posterior (under general conditions). In the face of only these asymptotic guarantees, it is actually quite miraculous that MCMC samplers \textit{can} often practically converge to the posterior in only a (relatively small) finite number of samples.

The predominant diagnostic method for determining whether practical convergence has occurred relies on the idea that the posterior distribution is the unique stationary distribution for an MCMC sampler. Therefore, it would appear that, if an MCMC sampling distribution stops changing, then convergence has occurred. Unfortunately, anyone who uses MCMC knows that the method is full of false dawns: the Markov chains can easily become stuck in areas of parameter space and observation over short intervals can mean that the sampling distribution \textit{appears} converged. Like furious bees trapped in a room of a house [cite video], MCMC samplers may fail to move because of narrow gaps joining neighbouring areas. With MCMC, absence of evidence of new areas of high posterior density is, time and again, not evidence of their absence.
